\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-02-24 trunk/arch.tex
updates to arch and results mostly cosmetic
final updates

\begin{figure}[!htb]
  \centering
  \subfloat[\small A \rx agent is composed of multiple reactors or
    control loops (rounded boxes) which are connected through state
    variables provided by one reactor ({\color{blue}blue} solid line)
    with multiple possible clients ({\color{green}green} dashed
    lines)]{\includegraphics[scale=0.45]{figs/AUV-agent.pdf}
  \label{fig:agent}
  }
  \qquad
  \subfloat[\small State information flow between the \texttt{Pilot}
  and the \texttt{Vehicle} reactors in an \rx instance.  Observations
  transit as a bottom-up flow and apply to the execution frontier
  $\tau$.  Goal information flow from the top to the bottom and
  relates to desired future values of state variables.]{
    \includegraphics[width=0.4\columnwidth]{figs/TREX_flow}
    \label{fig:trex:flow}
  }
\caption{\small \rx reactors and shared state variables.}
\end{figure}

@@

It is important to note that, in T-REX architecture, the only critical
task of a reactor is to successfully identify its state through
synchronization as the execution frontier advances. Indeed, a reactor
that fails to integrate its {\em external} state updates and/or
identify its {\em internal} state at any tick express that it is not
able anymore to explain the state of the world. In the agent point of
view that means that any state information from this reactor is not to
be trusted anymore and results on this reactor exclusion from the
execution cycle and consequent destruction. As the agent destroy 
the  faulty reactor is destroyed the agent produced on all of 
its former {\em internal} timelines the special observation 
\texttt{Failed} to all the reactors which did declare these state 
variables as {\em external} informing
them through this that this specific state variable is currently
unavailable. On the other hand all goals exchanged between reactors
are considered as rejectable by the reactor that manages them. As a
result the failure to find a plan for these goals, while to be avoided,
is not considered as a critical failure of the reactor and will only
result on these goals being ignored until new goals are received.

Take for example our AUV agent from Fig. \ref{fig:agent},
and assume that the \texttt{Sampling} reactor failed to synchronize;
in  such case this reactor is destroyed and both
\texttt{Sampling\_policy} and \texttt{Relevant\_data} state variables
produced the \texttt{Failed} observation. Therefore, the
\texttt{Survey controller} will received this observation on its {\em
  external} \texttt{Sampling\_policy} which indicates that this state
variable is not managed anymore. At this point depending on this
reactor implementation we have two possible outcomes:
\begin{enumerate}
\item \texttt{Survey  controller} model allows him to integrate this
  \texttt{Failed} observation. For example it can identify that, from
  now on, there's no way to take water samples. It still can 
  navigate as \texttt{Pilot} is available which allow him to continue
  this aspect of its current plan.
\item \texttt{Survey controller} models is not able to handle this
  unexpected failure resulting in turn on its failure to
  synchronize. This result on this reactor being also destroyed and
  the \texttt{Failed} observation is produced on the \texttt{survey}
  timeline.
\end{enumerate}
Even on the worst case scenario presented on this second case, we
note that a reactor failure only propagates up in the reactor graph. 
In our example, even if all the reactors that depends on
\texttt{Sampling} are entrained by its failure, we still keep the
three reactors \texttt{Vehicle}, \texttt{Pilot} and \texttt{Ship
  operation} allowing to  maintain the vehicle navigation control
(\texttt{Pilot}) along with high level safety operation 
(\texttt{Ship operation}). It illustrates how \rx can handle graceful
system degradation despite the failure of one or more reactors. It
also emphasize that it is critical to ensure that the low level
reactors are properly designed in order to be as robust as 
possible.

\begin{figure}[!htbp]
  \centering
  \vskip-1pc
  \includegraphics[width=0.55\columnwidth]{figs/tick-cycle}
  \caption{\small \rx reactor execution cycle: {\em Deliberation} is
    interrupted by {\em synchronization} at the beginning of every
    {\em tick} allowing integration of state information.}
  \label{fig:tick-exec}
  \vskip-0.8pc
\end{figure}

@@

@@

\item The \emph{synchronization solver} is a specialized \eu solver
  that integrates new information in the plan. This signifies the
  evolution of \emph{external} state variables and ensures that the
  reactor propagates these to identify current reactor state as well
  as to inform other reactors of any state change on its
  \emph{internal} timelines. This solver is summoned at the beginning
  of every single tick.

@@

\begin{figure}[!htb]
  \centering
  \subfloat[\small A plan produced before
  synchronization.]{\label{fig:shop:exec0}\includegraphics[width=0.7\columnwidth]{figs/shoping_exec_t0}} \\
  \subfloat[\small Result of synchronization after the {\em Going}
  observation was received by the reactor with the plan showed in
    \ref{fig:shop:exec0}.]{\label{fig:shop:exec1}\includegraphics[width=.7\columnwidth]{figs/shoping_exec_t1}}
    \caption{\small The Shopping example: Tokens in {\color{red}red}
      indicate observations, tokens in {\color{blue}blue} the goal
      that this plan attempts to solve for. Arrows indicate temporal
      constraints. }
  \label{fig:shop:exec}
\end{figure}

@@

\begin{figure}[!htbp]
  \centering
  \subfloat[Conflicting Observation]{\label{fig:shop:relax1}\includegraphics[width=0.7\columnwidth]{figs/shoping_exec_relax-1}}\\
  \subfloat[Relaxed Partial Plan]{\label{fig:shop:relax2}\includegraphics[width=0.7\columnwidth]{figs/shoping_exec_relax-2}}
  \caption{\small Illustration of a conflict during synchronization
    and the resulting relaxed plan for recovery.}
\end{figure}

>>>>

\begin{figure}[t]
  \centering
  \subfloat[\small A \rx agent is composed of multiple reactors or
    control loops (rounded boxes) which are connected through state
    variables provided by one reactor ({\color{blue}blue} solid line)
    with multiple possible clients ({\color{green}green} dashed
    lines).]{\includegraphics[scale=0.45]{figs/AUV-agent.pdf}
  \label{fig:agent}
  }
  \qquad
  \subfloat[\small State information flow between the \texttt{Pilot}
  and the \texttt{Vehicle} reactors in an \rx instance.  Observations
  transit as a bottom-up flow and apply to the execution frontier
  $\tau$.  Goal information flow from the top to the bottom and
  relates to desired future values of state variables.]{
    \includegraphics[width=0.4\columnwidth]{figs/TREX_flow}
    \label{fig:trex:flow}
  }
\caption{\small \rx reactors and shared state variables.}
\end{figure}

@@

\label{sec:delib-exec-as}
\begin{figure}[!htbp]
  \centering
  \vskip-1pc
  \includegraphics[width=0.55\columnwidth]{figs/tick-cycle}
  \caption{\small \rx reactor execution cycle: {\em Deliberation} is
    interrupted by {\em synchronization} at the beginning of every
    {\em tick} allowing integration of state information.}
  \label{fig:tick-exec}
  \vskip-0.8pc
\end{figure}

@@

\subsubsection{Handling Failure in \rx}
\label{sec:rx-reactor-failure}

The critical task of a \rx reactor is to successfully identify its
state through synchronization as the execution frontier advances. When
a reactor fails to integrate its {\em external} state update and/or
identify its {\em internal} state at any tick, it implies that it is
unable to explain the state of the world.  This in turn implies that
any state information from this reactor is suspect, resulting in the
exclusion of this reactor from the execution cycle and consequent
destruction. As the agent destroys the faulty reactor, a special
observation, \texttt{Failed}, is produced on all {\em internal}
timelines. Reactors which declare these very state variables as {\em
  external}, are informed of the failure of this timeline.  All goals
exchanged between reactors are considered \emph{rejectable} by the
reactor that manages them. As a result in such a situation, the
failure to find a plan for these goals, is \emph{not} considered a
critical failure of the reactor. It will only result on these goals
being ignored until newer goals are received.

Take for example our AUV agent from Fig. \ref{fig:agent}, and assume
that the \texttt{Sampling} reactor failed to synchronize. In such a
case this reactor is destroyed and both \texttt{Sampling\_policy} and
\texttt{Relevant\_data} timelines generate the \texttt{Failed}
observation as illustrated in Fig. \ref{fig:fail0}.  The
\texttt{Survey controller} will receive this \texttt{Failed}
observation on its {\em external} \texttt{Sampling\_policy} timeline
which indicates that this state variable is no longer managed by its
owner reactor. This leads to two potential outcomes based on how the
reactor is implemented:

\item The \texttt{Survey controller} model allows it to integrate the
  \texttt{Failed} observation. For example it can identify that,
  subsequent use of the water sampler is no longer possible.  The \rx
  agent can continue to navigate since \texttt{Pilot} is still
  available. In such a case the agent structure remains as in
  Fig. \ref{fig:fail0} with no consequent impact.

\item The \texttt{Survey controller} model is unable to handle this
  unexpected failure resulting in its failure to synchronize. This
  results in the \texttt{Survey controller} also failing, with
  subsequent destruction and flagging via the \texttt{Failed}
  observation on its \texttt{survey} timeline as shown in
  Fig. \ref{fig:fail1}.

\begin{figure}[b]
  \centering
  \subfloat[\small When the \texttt{Sampling} reactor fails to
  synchronize, it is removed from the reactor dependency graph along
  with all its {\em internal} timelines while generating a \texttt{Failed} observation]{\includegraphics[width=0.35\columnwidth]{figs/AUV-fail0.pdf}
  \label{fig:fail0}}
  \qquad
  \subfloat[\small In the worst case this \texttt{Failed} observation may imply
  other reactors that depend on the \texttt{Sampling} reactor might fail to
  synchronize leading to a bottom-up destruction of ``faulty'' reactors.]{
    \includegraphics[width=0.35\columnwidth]{figs/AUV-fail1}\label{fig:fail1}}
  \caption{\small Impact of the \texttt{Sampling} reactor failure in
    the AUV agent shown in Fig. \ref{fig:agent}}
\end{figure}

Even in the worst of the two cases presented above, we note that a
reactor failure only propagates \emph{up} in the reactor dependency
graph.  In our example, even if all the reactors that depend on
\texttt{Sampling} are entrained by its failure, \rx continues to keep
the reactors \texttt{Vehicle}, \texttt{Pilot} and \texttt{Ship
  operation} intact.  This ensures maintenance of vehicle control via
the \texttt{Pilot}, in addition to an abstract commanding mode for
safe operation via the \texttt{Ship operation} reactor, illustrating
how \rx can handle system degradation despite the failure of one or
more reactors, and doing so gracefully.

Consequently, by ensuring that lower level reactors closest to the
platform hardware are robustly designed, the \rx framework ensures
that system designers can focus their attention on critical components
while factoring in deliberative control of their platforms.

@@

\item The \emph{synchronization solver} is a specialized \eu solver
  that integrates new information in the plan. This signifies the
  evolution of \emph{external} timelines and ensures that the reactor
  propagates these to identify current reactor state as well as to
  inform other reactors of any state change on its \emph{internal}
  timelines. This solver is summoned at the beginning of every single
  tick.

@@

\begin{figure}[!b]
  \centering
  \subfloat[\small A plan produced before
  synchronization.]{\label{fig:shop:exec0}\includegraphics[width=0.7\columnwidth]{figs/shoping_exec_t0}} \\
  \subfloat[\small Result of synchronization after the {\em Going}
  observation was received by the reactor with the plan showed in
    \ref{fig:shop:exec0}.]{\label{fig:shop:exec1}\includegraphics[width=.7\columnwidth]{figs/shoping_exec_t1}}
    \caption{\small The Shopping example: Tokens in {\color{red}red}
      indicate observations, tokens in {\color{blue}blue} the goal
      that this plan attempts to solve for. Arrows indicate temporal
      constraints. }
  \label{fig:shop:exec}
\end{figure}

@@

\begin{figure}[!b]
  \centering
  \subfloat[Conflicting Observation]{\label{fig:shop:relax1}\includegraphics[width=0.7\columnwidth]{figs/shoping_exec_relax-1}}\\
  \subfloat[Relaxed Partial Plan]{\label{fig:shop:relax2}\includegraphics[width=0.7\columnwidth]{figs/shoping_exec_relax-2}}
  \caption{\small Illustration of a conflict during synchronization
    and the resulting relaxed plan for recovery.}
\end{figure}

### 2012-02-24 trunk/concepts.tex
final updates

  \begin{figure}[t]
    \centering \subfloat[\small Timelines representing initial state
    and goals for a Shopping
    Agent.]{\label{fig:europapr1}\includegraphics[scale=0.35]{figs/europa-pr-1.pdf}}\qquad
    \subfloat[\small State of timelines after satisfying a goal to be
    home by time $20$ on merging with initial
    state.]{\label{fig:europapr:merge}\includegraphics[scale=0.35]{figs/europa-pr-merge.pdf}}\qquad
    \subfloat[\small State of timelines after satisfying a goal to own
    a drill by time $10$ by activating corresponding
    token.]{\label{fig:europapr:activate}\includegraphics[scale=0.35]{figs/europa-pr-activate.pdf}}
    \caption{\small Evolution of a timeline from the Shopping Agent example}
  \end{figure}

@@

When a token is merged, all restrictions imposed on the merged token
are transferred to the active token upon which it is merged. Merging a
token requires finding a target active token that is compatible. Two
tokens are merge-compatible if they are instances of the same
predicate and there are no empty intersections between corresponding
token variables.  The effect of merging is illustrated in
Fig. \ref{fig:europapr:merge}

% \begin{figure} 
%   \centering
%   \includegraphics[scale=0.4]{figs/europa-pr-merge.pdf}
%   \caption{\small State of timelines after satisfying a goal to be
%     home by time $20$ on merging with initial state.}
%   \label{fig:europapr:merge}
% \end{figure}

The goal to be home by time $20$ is satisfied by merging with the
initial state that said that the Agent is home at time $0$.  The merge
is possible because the tokens are of the same type \texttt{At}, and
intersecting the domains for all of their variables doesn't yield an
empty domain (which represents an inconsistency), in this case the
token variables are the location, which is \texttt{Home} in both
cases, and the temporal variables (start, end, duration), which can be
intersected without causing an inconsistency. After the token
representing the goal is merged onto the token representing the
initial state, the resulting token represents not only the fact that
the Agent must be \texttt{At(Home)} by time 0 but also
\texttt{At(Home)} at time $20$, this is reflected in the new
constraints on the token's start and end variables as depicted in
Fig. \ref{fig:europapr:merge}.

When a token is activated, all domain rules associated with its type
are evaluated, which may lead to new subgoals being generated. In our
example, there could be a domain rule stating that if a Shopping Agent
doesn't already own a product, then it must \texttt{Buy} it before it
can \texttt{Own} it. In such a case, activating the
\texttt{Own(Drill)} token will cause a new \texttt{Buy(Drill)} to be
created in an \texttt{INACTIVE} state, which in turn could spawn its
own subgoals, like being at the location where the product can be
purchased before being able to buy it, and so on, until all domain
rules are satisfied.

% \begin{figure} \centering
%   \includegraphics[scale=0.4]{figs/europa-pr-activate.pdf}
%   \caption{\small State of timelines after satisfying a goal to own a
%     drill by time $10$ by activating corresponding token.}
% \label{fig:europapr:activate}
% \end{figure}

As Fig. \ref{fig:europapr:activate} shows the decision to merge the
goal to be \texttt{At(Home)} by time $20$ with the initial state was
not appropriate since the Agent will have to get away from
\texttt{Home} to be able to buy the drill which is inconsistent with
the constraints that resulted after the merge. As a result that
decision will have to be retracted at some point and the alternative
to activate or merge with a different token will have to be
considered.

@@

\begin{figure} \centering
  \includegraphics[scale=0.45]{figs/europa-pr-states.jpg}
  \caption{\small Possible transitions for a Token State}
\label{fig:europapr:states}
\end{figure}

@@

  \begin{figure} 
    \centering
    \includegraphics[scale=0.4]{figs/europa-pr-rules.pdf}
    \caption{\small Master and Slave tokens generated as a result of a domain rule.}
    \label{fig:europapr:rules}
  \end{figure}

>>>>

  \begin{figure}[t]
    \centering \subfloat[\small Timelines representing initial state
    and goals for a Shopping
    Agent.]{\label{fig:europapr1}\includegraphics[scale=0.35]{figs/europa-pr-1.pdf}}\qquad
    \subfloat[\small State of timelines after satisfying a goal to be
    home by time $20$ on merging with initial
    state.]{\label{fig:europapr:merge}\includegraphics[scale=0.35]{figs/europa-pr-merge.pdf}}\qquad
    \subfloat[\small State of timelines after satisfying a goal to own
    a drill by time $10$ by activating corresponding
    token.]{\label{fig:europapr:activate}\includegraphics[scale=0.35]{figs/europa-pr-activate.pdf}}
    \caption{\small Evolution of a timeline from the Shopping Agent example.}
  \end{figure}

@@

When a token is merged, all restrictions imposed on the merged token
are transferred to the active token upon which it is merged. Merging a
token requires finding a target active token that is compatible. Two
tokens are merge-compatible if they are instances of the same
predicate and there are no empty intersections between corresponding
token variables.  The effect of merging is illustrated in
Fig. \ref{fig:europapr:merge}. The goal to be home by time $20$ is
satisfied by merging with the initial state that said that the Agent
is home at time $0$.  The merge is possible because the tokens are of
the same type \texttt{At}, and intersecting the domains for all of
their variables doesn't yield an empty domain (which represents an
inconsistency), in this case the token variables are the location,
which is \texttt{Home} in both cases, and the temporal variables
(start, end, duration), which can be intersected without causing an
inconsistency. After the token representing the goal is merged onto
the token representing the initial state, the resulting token
represents not only the fact that the Agent must be \texttt{At(Home)}
by time 0 but also \texttt{At(Home)} at time $20$, this is reflected
in the new constraints on the token's start and end variables as
depicted in Fig. \ref{fig:europapr:merge}.

When a token is activated, all domain rules associated with its type
are evaluated, which may lead to new subgoals being generated. In our
example, there could be a domain rule stating that if a Shopping Agent
doesn't already own a product, then it must \texttt{Buy} it before it
can \texttt{Own} it. In such a case, activating the
\texttt{Own(Drill)} token will cause a new \texttt{Buy(Drill)} to be
created in an \texttt{INACTIVE} state, which in turn could spawn its
own subgoals, like being at the location where the product can be
purchased before being able to buy it, and so on, until all domain
rules are satisfied. As Fig. \ref{fig:europapr:activate} shows the
decision to merge the goal to be \texttt{At(Home)} by time $20$ with
the initial state was not appropriate since the Agent will have to get
away from \texttt{Home} to be able to buy the drill which is
inconsistent with the constraints that resulted after the merge. As a
result that decision will have to be retracted at some point and the
alternative to activate or merge with a different token will have to
be considered.

@@

\begin{figure} \centering
  \includegraphics[scale=0.45]{figs/europa-pr-states.jpg}
  \caption{\small Possible transitions for a token state.}
\label{fig:europapr:states}
\end{figure}

@@

  \begin{figure}[t]
    \centering
    \includegraphics[scale=0.4]{figs/europa-pr-rules.pdf}
    \caption{\small Master and Slave tokens generated as a result of a domain rule.}
    \label{fig:europapr:rules}
  \end{figure}

### 2012-02-24 trunk/europa-basics.tex
final updates

The next few sections will cover \eu in a bottom-up fashion, going
from the generic theoretical framework that underlies \eu, to the
specific representation, modeling and algorithms implemented by it.  A
brief introduction to Constraint Programming and Constraint-Based
Planning is provided followed by the specific representation that \eu
uses to undertake Constraint-Based Planning.  Subsequently, we provide
an overview of \eus architecture and how it can be embedded in a
planning application such as \rx. Finally, \eus modeling, inference
and search mechanisms are covered in detail so that the reader can
understand how \rx can perform deliberative planning and react to
execution events by embedding and customizing \eu.

@@

\begin{figure} \centering
  \includegraphics[scale=0.35]{figs/shopping_europa.pdf}
  \caption{\small A potential plan for a robot shopping for groceries.}
\label{fig:shop:shopping}
\vskip+0.1cm
\end{figure}

To articulate the fundamentals of automated planning briefly and use
that to motivate the mechanisms we use in our specific form of
planning we start with a simple example.

@@

At the core of the \rx framework, is the deliberation engine, \eu,
which has a rich legacy from NASA missions \cite{mus98,rajan00,
  jonsson00,aichang04, bresina05}. \eu is a versatile Constraint-based
Temporal Planner which continues to be deployed on a diverse set of
applications including recently, planning for the International Space
Station \cite{barreiro09}. We motivate this section with some problem
domains this planner can handle and then proceed to go into detail.

\paragraph{} {\em Constraint Satisfaction}: A canonical problem in
dealing with constraints is the $N$-Queens problem in which chess queens
must be placed on an  $N$x$N$ chessboard so no queens attack the other. % Fig.
% \ref{fig:nqueens-1} shows an example of a random positioning of queens
% on a  $N$x$N$ chessboard. Queens in violation of the non-attack constraint
% are highlighted in red.
% \begin{figure} \centering
%   \includegraphics[scale=0.35]{figs/Example-NQueens0.jpg}
%   \caption{\small N-Queens problem. Queens in violation of the
%     non-attack constraint are highlighted in red.}
% \label{fig:nqueens-1}
% \vskip+0.1cm
% \end{figure}
If we define $N$x$N$ variables Q$_{rc}$, $r \in [1,N]$, $c \in [1,N]$,
Q$_{rc} = 1$ if cell $r,c$ in the chessboard is occupied by a Queen,
$0$ otherwise. Then the following constraints need to be satisfied:

@@

\begin{figure}
\centering
\includegraphics[scale=0.35]{figs/Example-NQueens1.jpg}
\caption{\small N-Queens solution generated by \eu}
\label{fig:nqueens-2}
\vskip+0.1cm
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[scale=0.35]{figs/Example-UBO0.jpg}
\caption{\small A \eu solution to an \texttt{RCPSP} \cite{Bruckera99}
  problem.}
\label{fig:rcpsp-1}
\vskip+0.1cm
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[scale=0.35]{figs/Example-Shopping0.jpg}
\caption{\small A \eu solution to a shopping agent problem domain where
  the agent needs to buy Bananas, Milk and a Drill.}
\label{fig:shopping-1}
\vskip+0.1cm
\end{figure}

@@

A set of constraints and variables that describe a problem domain are
typically represented as a \emph{Constraint Network}, where the
variables are nodes and constraints are arcs between the variables. A
pair of variables $x \& y$ is said to be \emph{arc consistent} if for
every value in $x$'s domain, there is a value in $y$'s domain that is
consistent with all the constraints that connect $x$ and $y$.  A naive
approach to ensuring arc consistency consists of cycling through all
the variable pairs and performing constraint propagation until there
are no variable domain changes. The \texttt{AC-3} algorithm
\cite{mackworth77}, a popular implementation, improves over the naive
approach by ignoring constraints whose variables were not affected
since the last iteration. Fig. \ref{fig:constraintnet} shows a set of
constraints on $3$ variables ($x$, $y$ and $z$) represented as a
constraint network, and their domains before arc consistency is
enforced.  After the \texttt{AC-3} algorithm is executed, the domains
for the variables would be pruned to x=$\{2$\}, y=$\{2$\}, z=$\{2$\}.

\begin{figure}[!t]
\centering
\includegraphics[scale=0.35]{figs/constraint-net.pdf}
\caption{\small Variables and Constraints represented as a Constraint Network}
\label{fig:constraintnet}
\end{figure}

>>>>

The next few sections will cover \eu in a bottom-up fashion, going
from the generic theoretical framework that underlies \eue, to the
specific representation, modeling and algorithms implemented by it.  A
brief introduction to Constraint Programming and Constraint-Based
Planning is provided followed by the specific representation that \eu
uses to undertake Constraint-Based Planning.  Subsequently, we provide
an overview of \eus architecture and how it can be embedded in a
planning application such as \rxe. Finally, \eus modeling, inference
and search mechanisms are covered in detail so that the reader can
understand how \rx can perform deliberative planning and react to
execution events by embedding and customizing \eu.

@@

Planning as an activity is ingrained in human decision-making
\cite{berthoz} and its generalization into a computational framework
demonstrating machine intelligence long targeted by pioneers in
computer science \cite{computersthought}. Early work in this context
focused on logical theorem proving by computer \cite{green69} which
has formed the basis of much of the discipline of AI Planning. The
emphasis in these early attempts was to logically derive the existence
proof of a trajectory of actions for a hypothetical robot (moving
blocks, chess pieces, furniture etc) to achieve a specific goal given
specific initial conditions. Two important properties of
\emph{Soundness}\footnote{A planning algorithm is sound if invoked on
  a problem $P$ returns a plan which is a solution for $P$.} and
\emph{Completeness}\footnote{A planning algorithm is complete if
  invoked on a solvable problem $P$ is guaranteed to return a
  solution.} along with derived results for algorithmic complexity
\cite{corman} were a defining cornerstone for understanding how
different computational paradigms could actually be tractable first in
simulation and subsequently in actual robotic platforms. In this
chapter our emphasis is to focus on embedded AI Planning and the
execution of plans on an embedded robotic marine robotic
platform. While the platform is domain specific, the techniques we
highlight are general purpose as noted in Section \ref{sec:intro}.

\begin{figure}[b]
\centering
  \includegraphics[scale=0.35]{figs/shopping_europa.pdf}
  \caption{\small A potential plan for a robot shopping for groceries.}
\label{fig:shop:shopping}
\end{figure}

To articulate the fundamentals of automated planning briefly and use
that to motivate the mechanisms we use in our specific form of
planning we start with an example.

@@

At the core of the \rx framework, is the deliberation engine, \eue,
which has a rich legacy from NASA missions \cite{mus98,rajan00,
  jonsson00,aichang04, bresina05}. \eu is a versatile Constraint-based
Temporal Planner which continues to be deployed on a diverse set of
applications including recently, planning for the International Space
Station \cite{barreiro09}. We motivate this section with some problem
domains this planner can handle and then proceed to go into detail.

\paragraph{} {\em Constraint Satisfaction}: A canonical problem in
dealing with constraints is the $N$-Queens problem in which chess queens
must be placed on an  $N \times N$ chessboard so no queens attack the other. % Fig.
% \ref{fig:nqueens-1} shows an example of a random positioning of queens
% on a  $N$x$N$ chessboard. Queens in violation of the non-attack constraint
% are highlighted in red.
% \begin{figure} \centering
%   \includegraphics[scale=0.35]{figs/Example-NQueens0.jpg}
%   \caption{\small N-Queens problem. Queens in violation of the
%     non-attack constraint are highlighted in red.}
% \label{fig:nqueens-1}
% \vskip+0.1cm
% \end{figure}
If we define $N \times N$ variables Q$_{rc}$, $r \in [1,N]$, $c \in
[1,N]$, Q$_{rc} = 1$ if cell $r,c$ in the chessboard is occupied by a
Queen, $0$ otherwise. Then the following constraints need to be
satisfied:

@@

\begin{figure}
\centering
\includegraphics[scale=0.3]{figs/Example-NQueens1.jpg}
\caption{\small An $N$-Queens solution generated by \eu}
\label{fig:nqueens-2}
\end{figure}

@@

\begin{figure}[t]
\centering
\subfloat[\small A \eu solution to an \texttt{RCPSP} \cite{Bruckera99}
  problem.]{\label{fig:rcpsp-1}\includegraphics[scale=0.2]{figs/Example-UBO0.jpg}}\qquad
\subfloat[\small A \eu solution to a shopping agent problem domain where
  the agent needs to buy Bananas, Milk and a Drill.]{\label{fig:shopping-1}\includegraphics[scale=0.25]{figs/Example-Shopping0.jpg}}
\caption{\small \eu based solutions to scheduling and planning problems.}
\end{figure}

@@

@@

A set of constraints and variables that describe a problem domain are
typically represented as a \emph{Constraint Network}, where the
variables are nodes and constraints are arcs between the variables. A
pair of variables $x$ and $y$ is said to be \emph{arc consistent} if
for every value in $x$'s domain, there is a value in $y$'s domain that
is consistent with all the constraints that connect $x$ and $y$.  A
naive approach to ensuring arc consistency consists of cycling through
all the variable pairs and performing constraint propagation until
there are no variable domain changes. The \texttt{AC-3} algorithm
\cite{mackworth77}, a popular implementation, improves over the naive
approach by ignoring constraints whose variables were not affected
since the last iteration. Fig. \ref{fig:constraintnet} shows a set of
constraints on $3$ variables ($x$, $y$ and $z$) represented as a
constraint network, and their domains before arc consistency is
enforced.  After the \texttt{AC-3} algorithm is executed, the domains
for the variables would be pruned to x=$\{2$\}, y=$\{2$\}, z=$\{2$\}.

\begin{figure}[!t]
\centering
\includegraphics[scale=0.3]{figs/constraint-net.pdf}
\caption{\small Variables and Constraints represented as a Constraint Network}
\label{fig:constraintnet}
\end{figure}
